---
title: "Automatic ClustAssess pipeline. Generating the ClustAssess shiny-app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Automatic ClustAssess pipeline. Generating the ClustAssess shiny-app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
set.seed(1001)
```

```{r install, include = FALSE}
is_devtools <- require("devtools", quietly = TRUE)
if (!is_devtools) {
    install.packages("devtools", repos = "https://cloud.r-project.org")
    library(devtools)
}

is_seuratdata <- require("SeuratData", quietly = TRUE)
if (!is_seuratdata) {
    install_github('satijalab/seurat-data')
}
```

`ClustAssess` provide the option to run the entire stability pipeline automatically, by choosing the parameters based on their highest stability.
Another useful feature is the option to create, based on this object, a shiny app that user can interact with and perform the assessment of the PhenoGraph configuration and of the clusters.

```{r load}
library(Seurat)
library(ClustAssess)
library(SeuratData)
library(ggplot2)
packageVersion("ClustAssess") # should be 1.0.0
```

Process the PBMC 3k Seurat object similarly to the `Stability-based parameter assessment` vignette.

```{r}
InstallData("pbmc3k")
data("pbmc3k")
pbmc3k <- UpdateSeuratObject(pbmc3k)
pbmc3k <- PercentageFeatureSet(pbmc3k, pattern = "^MT-", col.name = "percent.mito")
pbmc3k <- PercentageFeatureSet(pbmc3k, pattern = "^RP[SL][[:digit:]]", col.name = "percent.rp")
# remove MT and RP genes
all.index <- seq_len(nrow(pbmc3k))
MT.index <- grep(pattern = "^MT-", x = rownames(pbmc3k), value = FALSE)
RP.index <- grep(pattern = "^RP[SL][[:digit:]]", x = rownames(pbmc3k), value = FALSE)
pbmc3k <- pbmc3k[!((all.index %in% MT.index) | (all.index %in% RP.index)), ]
pbmc3k <- subset(pbmc3k, nFeature_RNA < 2000 & nCount_RNA < 2500 & percent.mito < 7 & percent.rp > 7)
pbmc3k <- NormalizeData(pbmc3k, verbose = FALSE)
pbmc3k <- FindVariableFeatures(pbmc3k, selection.method = "vst", nfeatures = 3000, verbose = FALSE)

features <- dimnames(pbmc3k@assays$RNA)[[1]]
var_features <- pbmc3k@assays[["RNA"]]@var.features
n_abundant <- 3000
most_abundant_genes <- rownames(pbmc3k@assays$RNA)[order(Matrix::rowSums(pbmc3k@assays$RNA),
    decreasing = TRUE
)]

pbmc3k <- ScaleData(pbmc3k, features = features, verbose = FALSE)

pbmc3k <- RunPCA(pbmc3k,
    npcs = 30,
    approx = FALSE,
    verbose = FALSE,
    features = intersect(most_abundant_genes, pbmc3k@assays$RNA@var.features)
)
pbmc3k <- RunUMAP(pbmc3k,
    reduction = "pca",
    dims = 1:30,
    n.neighbors = 30,
    min.dist = 0.3,
    metric = "cosine",
    verbose = FALSE
)
```

Initialise the parallel backend.

```{r parallel_framwework}
RhpcBLASctl::blas_set_num_threads(1)
ncores <- 1
my_cluster <- parallel::makeCluster(
    ncores,
    type = "PSOCK"
)

doParallel::registerDoParallel(cl = my_cluster)
```

Define the feature sets of interest.

```{r feature_sets}
features <- dimnames(pbmc3k@assays$RNA)[[1]]
var_features <- pbmc3k@assays[["RNA"]]@var.features
n_abundant <- 3000
most_abundant_genes <- rownames(pbmc3k@assays$RNA)[order(Matrix::rowSums(pbmc3k@assays$RNA),
    decreasing = TRUE
)]

steps <- seq(from = 500, to = 3000, by = 500)
ma_hv_genes_intersection_sets <- sapply(steps, function(x) intersect(most_abundant_genes[1:x], var_features[1:x]))
ma_hv_genes_intersection <- Reduce(union, ma_hv_genes_intersection_sets)
ma_hv_steps <- sapply(ma_hv_genes_intersection_sets, length)
```

Apply the automatic assessment.

```{r automatic_run, warning = FALSE, cache = FALSE}
automm_output <- automatic_stability_assessment(
    expression_matrix = pbmc3k@assays$RNA@scale.data,
    n_repetitions = 10,
    n_neigh_sequence = seq(from = 5, to = 50, by = 5),
    resolution_sequence = seq(from = 0.1, to = 1, by = 0.1),
    features_sets = list(
        "HV" = var_features,
        "MA" = most_abundant_genes[1:3000]
    ),
    steps = list(
        "HV" = steps,
        "MA" = steps
    ),
    n_top_configs = 2,
    umap_arguments = list(
        min_dist = 0.3,
        n_neighbors = 30,
        metric = "cosine"
    ),
    save_temp = FALSE,
    verbose = TRUE
)
```

Close the connections opened when using multiple cores.
```{r remove_parallel}
foreach::registerDoSEQ()
```

Create the shiny app based on the ClustAssess output. You should also specify either a seurat object or a normalized expression matrix.
*Note*: Please make sure that the directory mentioned in the parameter `project_folder` is empty / doesn't exist.
```{r generate_app, eval = FALSE}
# generate using a seurat object
write_shiny_app(
    object = pbmc3k,
    assay_name = "RNA",
    clustassess_object = automm_output,
    project_folder = "clustassess_app_dir_seurat",
    shiny_app_title = "PBMC 3k dataset"
)

# generate using a normalized expression matrix
write_shiny_app(
    object = pbmc3k@assays$RNA@data,
    metadata = pbmc3k@meta.data,
    clustassess_object = automm_output,
    project_folder = "clustassess_app_dir_expr",
    shiny_app_title = "PBMC 3k dataset"
)
```

The app can be run using the following command.

```{r run_app, eval = FALSE}
shiny::runApp("clustassess_app_dir_seurat")
```

Session info
```{r ses_info}
sessionInfo()
```