---
title: "Flexibility of the dimensionality reduction assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flexibility of the dimensionality reduction assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r install}
is_seuratdata <- require("SeuratData", quietly = TRUE)
if (!is_seuratdata) {
    devtools::install_github('satijalab/seurat-data', upgrade = "never")
}

is_harmony <- require("harmony", quietly = TRUE)
if (!is_harmony) {
    install.packages("harmony", repos = "https://cloud.r-project.org")
}

is_rsamtool <- require("Rsamtools", quietly = TRUE)
if (!is_rsamtool) {
    is_biocmanager <- require("BiocManager", quietly = TRUE)
    if (!is_biocmanager) {
        install.packages("BiocManager", repos = "https://cloud.r-project.org")
        library(BiocManager)
    }

    install("Rsamtools")
}
```

```{r load}
library(Seurat)
library(SeuratData)
library(ClustAssess)
library(ggplot2)
library(harmony)
library(data.table)
library(Rsamtools)

n_repetitions <- 30
```

The `matrix processing` parameter of the `assess_feature_stability` function is a function that enables the user to specify any method to perform the dimensionality reduction prior to applying the UMAP algorithm and the clustering pipeline.
By default, the dimensionality reduction used in `ClustAssess` is a precise PCA using the `prcomp` package.
However, this function can be easily changed, as it will be shown in the following examples.

# ClustAssess using PCA

For the PCA example, we will use the PBMC 3k dataset from the `SeuratData` package. The preprocessing of the dataset is identical with the one performed in the stability pipeline vignette.
```{r process_pbmc_rna}
options(timeout=3600)
InstallData("pbmc3k")
data("pbmc3k")
pbmc3k <- UpdateSeuratObject(pbmc3k)
pbmc3k <- PercentageFeatureSet(pbmc3k, pattern = "^MT-", col.name = "percent.mito")
pbmc3k <- PercentageFeatureSet(pbmc3k, pattern = "^RP[SL][[:digit:]]", col.name = "percent.rp")
# remove MT and RP genes
all.index <- seq_len(nrow(pbmc3k))
MT.index <- grep(pattern = "^MT-", x = rownames(pbmc3k), value = FALSE)
RP.index <- grep(pattern = "^RP[SL][[:digit:]]", x = rownames(pbmc3k), value = FALSE)
pbmc3k <- pbmc3k[!((all.index %in% MT.index) | (all.index %in% RP.index)), ]
pbmc3k <- subset(pbmc3k, nFeature_RNA < 2000 & nCount_RNA < 2500 & percent.mito < 7 & percent.rp > 7)
pbmc3k <- NormalizeData(pbmc3k, verbose = FALSE)
pbmc3k <- FindVariableFeatures(pbmc3k, selection.method = "vst", nfeatures = 3000, verbose = FALSE)

features <- dimnames(pbmc3k@assays$RNA)[[1]]
var_features <- pbmc3k@assays[["RNA"]]@var.features
n_abundant <- 3000
most_abundant_genes <- rownames(pbmc3k@assays$RNA)[order(Matrix::rowSums(pbmc3k@assays$RNA),
    decreasing = TRUE
)]

pbmc3k <- ScaleData(pbmc3k, features = features, verbose = FALSE)
```

We notice that the `seurat_annotations` column has some missing values. For simplicity, we will replace them with "NA".
```{r mask_na}
mask <- is.na(pbmc3k$seurat_annotations)
pbmc3k$seurat_annotations <- as.character(pbmc3k$seurat_annotations)
pbmc3k$seurat_annotations[mask] <- "NA"
```

Select the features used for the stability assessment.
```{r select_features_rna}
features <- dimnames(pbmc3k@assays$RNA)[[1]]
var_features <- pbmc3k@assays[["RNA"]]@var.features
n_abundant <- 3000
most_abundant_genes <- rownames(pbmc3k@assays$RNA)[order(Matrix::rowSums(pbmc3k@assays$RNA),
    decreasing = TRUE
)]

steps <- seq(from = 500, to = 3000, by = 500)
ma_hv_genes_intersection_sets <- sapply(steps, function(x) intersect(most_abundant_genes[1:x], var_features[1:x]))
ma_hv_genes_intersection <- Reduce(union, ma_hv_genes_intersection_sets)
ma_hv_steps <- sapply(ma_hv_genes_intersection_sets, length)
```

Assess the stability of the dimensionality reduction when PCA is used as dimensionality reduction.
```{r feature_stability_pca, cache = FALSE}
matrix_processing_function <- function(dt_mtx, actual_npcs = 30) {
    actual_npcs <- min(actual_npcs, ncol(dt_mtx) %/% 2)

    RhpcBLASctl::blas_set_num_threads(foreach::getDoParWorkers())
    embedding <- stats::prcomp(x = dt_mtx, rank. = actual_npcs)$x

    RhpcBLASctl::blas_set_num_threads(1)
    rownames(embedding) <- rownames(dt_mtx)
    colnames(embedding) <- paste0("PC_", seq_len(actual_npcs))

    return(embedding)
}

pca_feature_stability <- assess_feature_stability(
    data_matrix = pbmc3k@assays[["RNA"]]@scale.data,
    feature_set = most_abundant_genes,
    resolution = seq(from = 0.1, to = 1, by = 0.1),
    steps = steps,
    n_repetitions = n_repetitions,
    feature_type = "MA",
    graph_reduction_type = "PCA",
    matrix_processing = matrix_processing_function,
    umap_arguments = list(
        min_dist = 0.3,
        n_neighbors = 30,
        metric = "cosine"
    ),
    ecs_thresh = 1,
    clustering_algorithm = 1
)
```

Plot the distribution of the celltypes on the UMAP embedding obtained on the top 1000 Most Abundant genes.
```{r umap_pca}
umap_df <- data.frame(pca_feature_stability$embedding_list$MA$"1000")
umap_df$celltypes <- pbmc3k$seurat_annotations
ggplot(umap_df, aes(x = UMAP_1, y = UMAP_2, color = celltypes)) +
    geom_point() +
    theme_classic()
```

# ClustAssess using Harmony

We can also modify the function by adding an addition post-processing step to the PCA. In this example, we will use the Harmony correction to remove the "batch effect" created by the celltypes.
*Note*: This example is meant to exemplify how to use the Harmony correction in the ClusAssess pipeline. The batch correction is actually not needed in the PBMC 3k dataset.

```{r feature_stability_harmony, cache = FALSE}
matrix_processing_function <- function(dt_mtx, actual_npcs = 30) {
    actual_npcs <- min(actual_npcs, ncol(dt_mtx) %/% 2)

    RhpcBLASctl::blas_set_num_threads(foreach::getDoParWorkers())
    embedding <- stats::prcomp(x = dt_mtx, rank. = actual_npcs)$x

    RhpcBLASctl::blas_set_num_threads(1)
    rownames(embedding) <- rownames(dt_mtx)
    colnames(embedding) <- paste0("PC_", seq_len(actual_npcs))

    embedding <- RunHarmony(embedding, pbmc3k$seurat_annotations, verbose = FALSE)

    return(embedding)
}

pca_harmony_feature_stability <- assess_feature_stability(
    data_matrix = pbmc3k@assays[["RNA"]]@scale.data,
    feature_set = most_abundant_genes,
    resolution = seq(from = 0.1, to = 1, by = 0.1),
    steps = steps,
    n_repetitions = n_repetitions,
    feature_type = "MA",
    graph_reduction_type = "PCA",
    matrix_processing = matrix_processing_function,
    umap_arguments = list(
        min_dist = 0.3,
        n_neighbors = 30,
        metric = "cosine"
    ),
    ecs_thresh = 1,
    clustering_algorithm = 1,
    verbose = TRUE
)
```

Plot the distribution of the celltypes on the UMAP embedding obtained on the top 1000 Most Abundant genes.
```{r umap_harmony}
umap_df <- data.frame(pca_harmony_feature_stability$embedding_list$MA$"1000")
umap_df$celltypes <- pbmc3k$seurat_annotations
ggplot(umap_df, aes(x = UMAP_1, y = UMAP_2, color = celltypes)) +
    geom_point() +
    theme_classic()
```


# ClustAssess in the scATAC-seq data

In this example we will showcase the flexibility of the `assess_feature_stability` function by using the ATAC-seq data.
For this example, we will use the multiome PBMC dataset from the `SeuratData` package.
```{r load_atac}
library(Signac)
InstallData("pbmcMultiome")
data("pbmc.atac")
```

As presented in the (Signac)(https://stuartlab.org/signac/articles/pbmc_vignette) package, the ATAC-seq data is usually processed using the TF-IDF normalization followed by the the calculation of the singular values. These two steps are also known as LSI (Latent Semantic Indexing).
```{r preprocess_atac}
pbmc.atac <- RunTFIDF(pbmc.atac)
```

Identify the highly variable peaks.
```{r features_atac}
pbmc.atac <- FindTopFeatures(pbmc.atac, min.cutoff = "q5")
var_peaks <- pbmc.atac@assays$ATAC@var.features[seq_len(3000)]
```

To speedup the assessment, set a parallel backend with 6 cores.

```{r parallel_backend}
RhpcBLASctl::blas_set_num_threads(1)
ncores <- 1
if (ncores > 1) {
    my_cluster <- parallel::makeCluster(
        ncores,
        type = "PSOCK"
    )

    doParallel::registerDoParallel(cl = my_cluster)
}
```
Assess the stability of the dimensionality reduction by varying the number of highly variable peaks.
```{r feature_stability_atac, cache = FALSE}
matrix_processing_function <- function(dt_mtx, actual_n_singular_values = 50) {
    actual_n_singular_values <- min(actual_n_singular_values, ncol(dt_mtx) %/% 2)

    RhpcBLASctl::blas_set_num_threads(foreach::getDoParWorkers())
    embedding <- RunSVD(Matrix::t(dt_mtx), n = actual_n_singular_values, verbose = FALSE)@cell.embeddings
    # remove the first component, as it does contain noise - see the Signac vignette
    embedding <- embedding[, 2:actual_n_singular_values]

    RhpcBLASctl::blas_set_num_threads(1)
    rownames(embedding) <- rownames(dt_mtx)
    colnames(embedding) <- paste0("LSI_", seq_len(actual_n_singular_values - 1))

    return(embedding)
}

lsi_atac_feature_stability <- assess_feature_stability(
    data_matrix = pbmc.atac@assays[["ATAC"]]@data,
    feature_set = var_peaks,
    resolution = seq(from = 0.1, to = 1, by = 0.1),
    steps = steps,
    n_repetitions = n_repetitions,
    feature_type = "HV_peaks",
    graph_reduction_type = "PCA",
    matrix_processing = matrix_processing_function,
    umap_arguments = list(
        min_dist = 0.3,
        n_neighbors = 30,
        metric = "cosine"
    ),
    ecs_thresh = 1,
    clustering_algorithm = 1,
    verbose = TRUE
)
foreach::registerDoSEQ()
```

Plot the distribution of the celltypes on the UMAP embedding obtained on the top 1000 Highly Variable peaks.
```{r umap_atac}
umap_df <- data.frame(lsi_atac_feature_stability$embedding_list$HV_peaks$"1000")
umap_df$celltypes <- pbmc.atac$seurat_annotations
ggplot(umap_df, aes(x = UMAP_1, y = UMAP_2, color = celltypes)) +
    geom_point() +
    theme_classic()
```

# Session info
```{r session_info}
sessionInfo()
```
