---
title: "Stability-based parameter assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stability-based parameter assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(1001)
```


```{r setup, warning = F, echo = F, message = F}
library(Seurat)
library(readr)
library(Matrix)
library(ggplot2)
library(ClustAssess)
library(patchwork)
```

In this vignette we illustrate a data-driven pipeline for the assessment of optimal clustering parameters for a given dataset. The case study is based on a subset of cells from the Cuomo et al 2020 (Single-cell RNA-sequencing of differentiating iPS cells reveals dynamic genetic effects on gene expression [ref](https://www.nature.com/articles/s41467-020-14457-z)) dataset. Specifically, 6 donors and 4 time points were chosen, comprising 1880 cells and 35430 expressed genes. The input is provided as a 35430 x 1880 expression level matrix containing the raw quantification (in csv format).

```{r load}
cell_metadata = read.csv('http://bioinf.stemcells.cam.ac.uk:3838/clustassess_1gmg5bq/cuomo_metadata.csv', sep=',', header = TRUE, row.names = 1)
cell_metadata$donor = as.factor(cell_metadata$donor)
cell_metadata$day = as.factor(cell_metadata$day)

cuomo.counts = read_csv('http://bioinf.stemcells.cam.ac.uk:3838/clustassess_1gmg5bq/all_genes_with_RP_MT.csv.gz')
gene_names = cuomo.counts$gene_name
cuomo.counts = cuomo.counts[, 1:1880]

cuomo =  CreateSeuratObject(cuomo.counts, row.names = gene_names)

cuomo = PercentageFeatureSet(cuomo, pattern = '^MT-', col.name = 'percent.mito')
cuomo = PercentageFeatureSet(cuomo, pattern = "^RP[SL][[:digit:]]", col.name = 'percent.rp')
one_level = as.factor(rep("cuomo", dim(cuomo)[2]))
  
cuomo@meta.data$level.ident = one_level
cuomo@meta.data$day.ident = cell_metadata$day
cuomo@meta.data$donor.ident = cell_metadata$donor
Idents(cuomo) = "donor.ident"

rm(cuomo.counts)
gc()
```


The analysis is based on the Seurat pipeline (Hao et al 2021, Integrated analysis of multimodal single-cell data, [ref](https://www.sciencedirect.com/science/article/pii/S0092867421005833?via%3Dihub)); subsequent to the filtering step, the mitochondrial and ribosomal genes are excluded from the downstream analysis.

```{r remove_mt_rp}
# remove MT and RP genes
all.index = 1:nrow(cuomo)
MT.index <- grep(pattern = "^MT-", x = rownames(cuomo), value = FALSE)
RP.index = grep(pattern = "^RP[SL][[:digit:]]", x = rownames(cuomo), value = FALSE)
cuomo = cuomo[!((all.index %in% MT.index) | (all.index %in% RP.index)  ), ]
```

The data is normalized and scaled using the Seurat's `NormalizeData` and `ScaleData` methods. The variable feature set is obtained by running the `FindVariableFeatures` function from the same package.

```{r normalize, warning=F}
cuomo = NormalizeData(cuomo, verbose = F)
cuomo = FindVariableFeatures(cuomo, selection.method = "vst", nfeatures = 3000, verbose = F)

features = dimnames(cuomo@assays$RNA)[[1]]
var_features = cuomo@assays[["RNA"]]@var.features

cuomo = ScaleData(cuomo, features = features, verbose = F)

cuomo = RunPCA(cuomo, npcs = 30, approx = F, verbose = F, features = intersect(rownames(cuomo@assays$RNA)[order(Matrix::rowSums(cuomo@assays$RNA), decreasing=TRUE)][1:3000], cuomo@assays$RNA@var.features))
cuomo = RunUMAP(cuomo,reduction = "pca", dims = 1:30, n.neighbors = 30, min.dist = 0.3, metric = "cosine", verbose = F)
raw_umap = cuomo@reductions$umap@cell.embeddings
```

# QC

QC overview for the Cuomo dataset.

(A) time point induced separation of cells;
(B) distributions of cells per donor, we notice a good overlap across all 6 donors, supporting the assumption of no significant batch effect.
(C) –(F) distributions of sequencing depths (C), number of features (D), and percentages of reads incident to mitochondrial genes (E) and ribosomal genes (F) illustrated using the colour gradient.

No significant effects are noted.

```{r qc, fig.width = 7, fig.height=7, include = T, echo = F}
wrap_plots(lapply(c(DimPlot(cuomo, group.by=c('day.ident', 'donor.ident'), combine = F),
                    FeaturePlot(cuomo, features = c('nFeature_RNA', 'nCount_RNA','percent.mito', 'percent.rp'), combine = F)), 
                  function(x) { x + labs(title = element_blank()) + theme(legend.key.size = unit(0.3, "cm"),
                                                                          legend.key.width = unit(0.1,"cm"),
                                                                          legend.text = element_text(size = 7),
                                                                          axis.text = element_text(size = 7),
                                                                          axis.title = element_text(size = 9)) })) + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 3)
```

*Note*: The following sections provide an example of applying the ClustAssess summary assessment methods. The values of the parameters should be varied and critically assessed for each individual dataset. Also, we recommend increasing the number of repetitions used for stability inference, to ensure reliable and robust conclusions. 

```{r nrep}
n_repetitions = 30
```

The following analysis is performed by varying the seed multiple times. The `n_cores` user-defined parameter indicates the number of parallel threads used to run the pipeline. Setting the value of 1 will result in a sequential run of all iterations. We encourage the usage of a higher number of cores when possible, as it will decrease the runtime of the assessment pipeline. Please note that the number of cores should be chosen carefully, as each newly created process will add to the memory usage.

```{r ncores}
n_cores = 1
```

# Dimensionality reduction

The first parameter that we will evaluate and note to substantially influence the dimensionality reduction (linear or non-linear) is the feature set. The Seurat pipeline relies on the highly variable genes (default) for the computation of the Principal Components; the Monocle pipeline relies on all the features. The top x most variable genes or most abundant genes can be selected, where x varies according to the characteristics of the dataset; frequently used values are 500, 1000, 2000, 3000 etc.

To showcase the importance of this parameter, we assess three different sets of features:

* **highly variable genes**: this set can be obtained in the Seurat pipeline by using either the `SCTransform` or `FindVariableFeatures` method. Depending on the method, using the default parameters will get a number of 2000 or 3000 genes.
* **most abundant genes**: this set can be obtained by sorting the genes based on their expression level.
* **intersection between highly variable and most abundant genes**: this set is obtained by intersecting a subset of the most abundant genes and the whole highly variable feature set. Taking the intersection will lead to a smaller number of genes compared to the initial sets.

```{r feature_sets}
most_abundant_genes = rownames(cuomo@assays$RNA)[order(Matrix::rowSums(cuomo@assays$RNA), decreasing=TRUE)]
ma_hv_genes_intersection = intersect(most_abundant_genes[1:3000], var_features)
steps = seq(from = 500, to = 3000, by = 500)
ma_hv_steps = sapply(steps, function(x) { length(intersect(most_abundant_genes[1:x], var_features))})
```

`get_feature_stability` is a function that explores the stability of a feature set. The input consists of a normalized expression matrix, that is summarized using either PCA or UMAP. The obtained embedding is then processed using the Leiden community detection method (on default values as described in the Seurat pipeline). This process is iterated across random seeds. The method outputs a list containing an UMAP embedding, the cell assignment corresponding to the most frequent partition and the Element-Centric Consistency (ECC) across all communities obtained throughout the iterations. The higher the Element-Centric Consistency, the more stable the clustering.

```{r, include = FALSE}
start_time_feature_stability = Sys.time()
```


```{r feature_stability, warning=FALSE}
pca_feature_stability_object = c(get_feature_stability(data_matrix = cuomo@assays[["RNA"]]@scale.data,
                                                       feature_set = most_abundant_genes,
                                                       steps = steps,
                                                       n_repetitions = n_repetitions,
                                                       feature_type = "MA",
                                                       graph_reduction_type = "PCA",
                                                       npcs = 30,
                                                       min_dist = 0.3,
                                                       n_neighbors = 30,
                                                       metric = "cosine",
                                                       ncores = n_cores,
                                                       ecs_thresh = 1,
                                                       algorithm = 1),
                                 get_feature_stability(data_matrix = cuomo@assays[["RNA"]]@scale.data,
                                                       feature_set = var_features,
                                                       steps = steps,
                                                       n_repetitions = n_repetitions,
                                                       feature_type = "HV",
                                                       graph_reduction_type = "PCA",
                                                       npcs = 30,
                                                       min_dist = 0.3,
                                                       n_neighbors = 30,
                                                       metric = "cosine",
                                                       ncores = n_cores,
                                                       ecs_thresh = 1,
                                                       algorithm = 1),
                                 get_feature_stability(data_matrix = cuomo@assays[["RNA"]]@scale.data,
                                                       feature_set = ma_hv_genes_intersection,
                                                       steps = ma_hv_steps,
                                                       n_repetitions = n_repetitions,
                                                       feature_type = "MA_HV",
                                                       graph_reduction_type = "PCA",
                                                       npcs = 30,
                                                       min_dist = 0.3,
                                                       n_neighbors = 30,
                                                       metric = "cosine",
                                                       ncores = n_cores,
                                                       ecs_thresh = 1,
                                                       algorithm = 1))
```

The function `get_feature_stability` outputs the stability of different feature sets and concatenates the results to ease cross-comparisons. The plot illustrates the stability assessment, across 30 runs, each with different random seeds, on the three gene sets described above. We assessed the stability of each set on incremental number of selected genes. Thus, the boxplots are arranged in groups of three (the number of evaluated feature sets); above each boxplot we specify the number of elements per subset. For the Cuomo case study, a conclusion that is supported based on this plot is that the most abundant set is more stable than the other two options, since it has higher Element-Centric Consistency scores.

```{r feature_stability_boxplot, fig.width = 7, fig.height=5}
plot_feature_stability_boxplot(pca_feature_stability_object, text_size  = 2.5) + 
  theme(legend.position = c(1,0),
        legend.justification = c(1,0))
```

Another angle for assessing the stability is centred on the comparison between consecutive steps, for each feature set, performed using the Element-Centric Similarity on the most frequent partitions from each step. The aim is evaluating the effect of increasing the number of genes on the final partitions, and, indirectly, determining the transition from the signal to the noise zone. For the Cuomo case study, we observe an increase in similarity between consecutive steps for the most abundant and highly variable genes, suggesting that selecting more genes would lead to a more robust partitioning.

```{r feature_stability_ecs_incremental, fig.width=7, fig.height=5}
plot_feature_stability_ecs_incremental(pca_feature_stability_object, dodge_width = 1, text_size = 2) +
    theme(legend.position = c(1,0),
        legend.justification = c(1,0))
```

To enhance the summarized information presented as boxplots, we use `plot_feature_stability_mb_facet` to visualize the corresponding facet plot displaying the most frequent partitions for each feature set and step on an UMAP embedding. This plot reveals additional insights i.e. the stability of the most abundant genes can be explained by the scattered distribution of the cells across multiple islands. The highly variable and the intersection gene sets tend to produce more compact partitions, that may be linked to biologically relevant cell types.

```{r feature_stability_mb_facet, fig.width = 7, fig.height = 15}
plot_feature_stability_mb_facet(pca_feature_stability_object, text_size = 3)
```

To further understand the areas of instability, we provide an additional plot where we display, on the same UMAPs as before, the Element-Centric Consistency score of the partitions obtained for each step and feature set. For the Highly Variable and the intersection sets, we note that the instability appears in groups of cells with high density. These visualizations can help us understand the relationship between stability and the topology of the data.

```{r feature_stability_ecs_facet, fig.width = 7, fig.height = 15}
plot_feature_stability_ecs_facet(pca_feature_stability_object)
```

```{r, include = F}
stop_time_feature_stability = Sys.time()
```


For the Cuomo case study, based on the plots presented above, we conclude that the most abundant gene set is not a suitable choice, as its stability is explained by the presence of multiple isolated groups of cells. Thus, we settle on using the highly variable genes set, as it produces a compact distribution of cells that is consistent across different seeds.

```{r update_object}
cuomo = cuomo[var_features, ]

cuomo = RunPCA(cuomo, npcs = 30, approx = F, verbose = F)
cuomo = RunUMAP(cuomo,reduction = "pca", dims = 1:30, n.neighbors = 30, min.dist = 0.3, metric = "cosine", verbose = F)
```


# Graph construction

The next step in a standard single-cell analysis pipeline is building the graph using the nearest neighbour algorithm. The following parameters influence the final partitioning:

* **base embedding**: the graph can be built on either the PCA or the UMAP embedding (using the expression matrix isn’t recommended, as the distances would be noisier and the runtime would increase)
* **the number of neighbours**
* **the graph type**: the graph can be either unweighted (NN case) or based on a weighted Shared-Nearest Neighbours (SNN) graph. For the latter, the weights are computed using the Jaccard Similarity Index (JSI) between the neighbourhoods of two cells.


`get_nn_conn_comps` is a method used to link the number of neighbours and the number of connected components (a connected component is a subgraph within which there exists a path between every pair of nodes). The method accepts as input both PCA and UMAP reductions; the distribution of the number of connected components obtainable for the given number of neighbours is calculated across random seeds. The output is a list containing, for each number of neighbours, an array with the number of connected components obtained on different seeds. As before, the objects can be concatenated to facilitate comparisons between different configurations and reductions.

```{r, include = FALSE}
start_time_nn_conn = Sys.time()
```


```{r nn_conn_comps}
nn_conn_comps_object = c(get_nn_conn_comps(object = cuomo@reductions$pca@cell.embeddings,
                                           n_neigh_sequence = c(c(1,2,3,4), seq(from = 5, to = 30, by = 5)),
                                           n_repetitions = n_repetitions,
                                           graph_reduction_type = "UMAP",
                                           ncores = n_cores,
                                           min_dist = 0.3,
                                           n_neighbors = 30,
                                           metric = "cosine"),
                         get_nn_conn_comps(object = cuomo@assays[["RNA"]]@scale.data,
                                           n_neigh_sequence = c(c(1,2,3,4), seq(from = 5, to = 30, by = 5)),
                                           n_repetitions = n_repetitions,
                                           graph_reduction_type = "PCA",
                                           ncores = n_cores,
                                           nv = 30))
```

The following plot describes the covariation between the number neighbours and the number of connected components obtained using both PCA and UMAP reductions as base for graph building. As the number of neighbours increases, the number of connected components decreases (this is an expected result, as increasing the number of neighbours result in a better connected graph). Please note that the number of connected components provides a lower bound on the number of clusters we can obtain by downstream community detection algorithms such as Louvain and Leiden.

Another comparison may be performed on the PCA and UMAP embeddings to assess graph connectivity. We note that, when using UMAPs, a better separation of the cells is achieved; this is not observed for PCAs, for which a quick convergence to a connected graph (which is a graph where there exists a path between every two nodes) is obtained. The link between the number of connected components and the functional interpretation of the results is that the former could be interpreted as the minimum number of clusters that could be obtained with the given number of neighbours e.g. for the UMAP case, to obtain 5 communities from the clustering step, we would require at least 10 nearest neighbours.

```{r connected_comps_evolution, fig.width=7, fig.height=5}
plot_connected_comps_evolution(nn_conn_comps_object)
```

```{r, include = F}
stop_time_nn_conn = Sys.time()
```

`get_nn_importance` is a method used for assessing the impact of number of nearest neighbours, of the graph type (NN or SNN) and of the base embedding (PCA or UMAP) in determining the final number of clusters. The input is either a normalized expression matrix or a PCA embedding, a range for the number of neighbours and the reduction type used for building the graph. The graph type is not provided, as the method will check both cases i.e. aaa and bbb. The output comprises a list of configurations (reduction type, and the graph type). Each configuration will contain a list of unique partitions obtained from multiple runs (iterations) for each selected value of the nearest neighbours.

```{r, include = FALSE}
start_time_nn_importance = Sys.time()
```


```{r nn_importance}
nn_importance_object = mapply(c,
                              get_nn_importance(object = cuomo@assays[["RNA"]]@scale.data,
                                         n_neigh_sequence = seq(from = 5, to = 30, by = 5),
                                         n_repetitions = n_repetitions,
                                         graph_reduction_type = "PCA",
                                         ecs_thresh = 1,
                                         ncores = n_cores,
                                         algorithm = 1,
                                         nv = 30),
                              get_nn_importance(object = cuomo@reductions$pca@cell.embeddings,
                                         n_neigh_sequence = seq(from = 5, to = 30, by = 5),
                                         n_repetitions = n_repetitions,
                                         graph_reduction_type = "UMAP",
                                         transpose = TRUE,
                                         ecs_thresh = 1,
                                         ncores = n_cores,
                                         algorithm = 1,
                                         min_dist = 0.3,
                                         n_neighbors = 30,
                                         metric = "cosine"),
                              SIMPLIFY = FALSE
)
```

We input the resulting object into `plot_n_neigh_k_correspondence`, to generate the following plot; it summarizes the relationship between the number of nearest neighbours and k, the number of clusters. As for the # neighbours vs # connected components correspondence plot, we note a similar descending trend of the number of clusters proportional to the increase the number of nearest neighbours. The plot also illustrates the difference between the two graph types: SNN reaches more stable results in terms of the number of clusters (over multiple iterations). If the initial object contains graphs based on both UMAP and PCA embedding, this plot will showcase the impact of this choice, as well.

```{r n_neigh_k_correspondence, fig.width = 7, fig.height=5}
plot_n_neigh_k_correspondence(nn_importance_object)
```

The stability of these parameters can be also evaluated using the Element-Centric Consistency applied on the partition list obtained over the multiple runs. The following summary plot underlines the evolution of the consistency as the number of neighbours increases; the SNN graph yields stable results even for smaller number of nearest neighbours, whereas the NN exhibits a stability improvement as the number of neighbours increases.

```{r n_neigh_ecs, fig.width=7, fig.height=5}
plot_n_neigh_ecs(nn_importance_object)
```

```{r, include = F}
stop_time_nn_importance = Sys.time()
```

# Graph clustering

The final step in a standard single-cell analysis pipeline is applying a graph-based clustering method. Choosing a community detection algorithm has a significant impact on the partitioning results. Using the `get_clustering_difference` we assess the stability and reproducibility of results obtained using various graph clustering methods available in the Seurat package: Louvain, Louvain refined, SLM and Leiden.

```{r clustering_diff}
cuomo = RunPCA(cuomo, npcs = 30, approx = F, verbose = F)
cuomo = RunUMAP(cuomo,reduction = "pca", dims = 1:30, n.neighbors = 30, min.dist = 0.3, metric = "cosine", verbose = F)
adj_matrix = FindNeighbors(cuomo@reductions$umap@cell.embeddings, k.param = 25, nn.method = "rann", verbose = F)$snn

start_time_clustering_method = Sys.time()
clustering_diff_obj = get_clustering_difference(graph_adjacency_matrix = adj_matrix,
                                                resolution = seq(from = 0.5, to = 1, by = 0.1),
                                                n_repetitions = n_repetitions,
                                                ecs_thresh = 1,
                                                ncores = n_cores,
                                                algorithm = 1:3)
```

The stability of the methods is evaluated using the Element-Centric Consistency (ECC), which is applied on the partition list obtained over multiple runs. The following summary plot underlines the variation of the consistency as the resolution parameter increases. Above each boxplot we display the number of clusters corresponding to the most frequent partition for the clustering method with the specific resolution parameter. Increasing the resolution leads to more clusters; in addition, we note similar outputs for all three algorithms. We also notice that Louvain and its refined version are less consistent for higher values of the resolution parameter. Thus, SLM is a more stable choice when it comes to clustering.

```{r clustering_difference_boxplot, fig.width = 7, fig.height=5}
plot_clustering_difference_boxplot(clustering_diff_obj)
```

To further underline the areas of instability, we provide an additional plot where we display, on the same UMAPs as before, the ECC score of the partitions obtained for each resolution parameter with each clustering method. We notice how increasing the resolution is causing a change of the unstable areas. For example, when resolution is 0.5 or 0.6, all algorithms are unstable in areas from the bottom-left islands. In the case where the resolution is 1, the area of instability has moved to the top-left island.

```{r clustering_difference_facet, fig.width= 7, fig.height=10}
plot_clustering_difference_facet(clustering_diff_obj, cuomo@reductions$umap@cell.embeddings)
```

```{r, include = F}
stop_time_clustering_method = Sys.time()
```


```{r, include = F}
start_time_resolution = Sys.time()
```

As previously suggested, the resolution parameter affects the number of clusters obtained. The `get_resolution_importance` evaluates the effect of a range of values for the resolution parameter on the stability of the output. Besides the resolution, the user can input multiple values of the following parameters: number of neighbours, the graph type and the clustering method. The function will return a list associated with each parameter value combination. For each resolution value, the list will contain multiple sublists of partitions corresponding to specific number of clusters.

```{r resolution_importance}
resolution_gridsearch = get_resolution_importance(embedding = cuomo@reductions$umap@cell.embeddings,
                                                  resolution = seq(from = 0.5, to = 1, by = 0.1),
                                                  n_neigh = 30,
                                                  n_repetitions = n_repetitions,
                                                  clustering_method = c(1,2),
                                                  graph_type = 2,
                                                  ecs_thresh = 1,
                                                  ncores = n_cores)
```

The resulting object can be visualized with `plot_k_resolution_corresp` as shown in the code below; we showcase the relationship between the number of clusters and the resolution value. This plot also provides information on suitable resolution values for predefined number of clusters. The colour gradient is proportional to the number of occurrences of the most frequent partition. It can also be used as proxy to describe the co-variation between (k, resolution). Lighter (higher) values indicate that little variation is observed, on changes of the random seed, between the number of clusters and the resolution value.

```{r k_resolution_corresp, fig.width = 7, fig.height=5}
plot_k_resolution_corresp(resolution_gridsearch) + ggtitle("resolution - k correspondence with ecs threshold = 1")
```

By default, the `get_resolution_importance` function (as well as the previous functions part of the assessment pipeline) relies on the merging of identical partitions. Two partitions are identical if their contingency table form a square diagonal matrix. Merging is also applied to reduce the object size and to minimize the runtime for calculating the Element-Centric Consistency score.
In practice, we also noticed minor differences between two partitions (only 5 cells change their cluster labels). For the merging of identical partitions, these would be classed as different, although, in practice, the biological interpretation is likely to be the same. To address this we included a similarity threshold for additional flexibility i.e.: two partitions are considered identical and will be merged if the Element-Centric Similarity Score between them is above a user-defined ECS threshold.

```{r merge_partitions}
resolution_gridsearch_thresh_99 = merge_partitions(resolution_gridsearch,
                                                   ecs_thresh = 0.99,
                                                   ncores = n_cores)
```

Lowering the ECS threshold leads, in general, to higher number of occurences for the most frequent partition.

```{r k_resolution_corresp_merged, fig.width = 7, fig.height=5}
plot_k_resolution_corresp(resolution_gridsearch_thresh_99) + ggtitle("resolution - k correspondence with ecs threshold = 0.99")
```

The `merge_resolutions` function diminishes the dependency between the resolution and the numver of clusters (k) i.e. the partitions with the same number of clusters, part of the same list for each configuration, are merged.

```{r merge_resolutions}
k_parts_obj = lapply(resolution_gridsearch, function(config_name) {
   merge_resolutions(config_name,
                     ncores = n_cores)
})
```

The following plot showcases the co-variation between the stability of the number of clusters and the number of different partitions resulting from changes on the seed or the resolution parameter. A high number of different partitions indicates a high stability for a given number of clusters. The colour gradient is proportional to the frequency of the most common partition with a given number of clusters. The colour gradient is proportional to the frequency of the most common partition with a given number of clusters. We note that, even if we obtain a high number of different partitions, if the frequency of the most common one is close to 1, then the overall stability is high. Observing a high number of partitions, each with low frequency, indicates high instability.

```{r k_n_partitions, fig.width=7, fig.height=5}
plot_k_n_partitions(k_parts_obj) + ggtitle("k - # partitions correspondence with ecs threshold = 1")


k_parts_obj_thresh_99 = merge_partitions(k_parts_obj,
                                         ecs_thresh = 0.99,
                                         ncores = n_cores)

plot_k_n_partitions(k_parts_obj_thresh_99) + ggtitle("k - # partitions correspondence with ecs threshold = 0.99")
```

```{r, include = F}
stop_time_resolution = Sys.time()
```

We conclude by summarizing the runtime for executing each step of the assessment pipeline. The results observed on other computers may vary; the runtime is also affected by the size of the date (i.e number of cells and features), the data itself and by actual values for the various parameters.

```{r}
paste("Feature stability methods runtime:", format(as.numeric(stop_time_feature_stability - start_time_feature_stability, units = "mins")), "minutes")
paste("NN - # connected components methods runtime:", format(as.numeric(stop_time_nn_conn - start_time_nn_conn, units = "mins")), "minutes")
paste("NN importance methods runtime:", format(as.numeric(stop_time_nn_importance - start_time_nn_importance, units = "mins")), "minutes")
paste("Clustering importance methods runtime:", format(as.numeric(stop_time_clustering_method - start_time_clustering_method, units = "mins")), "minutes")
paste("Resolution gridsearch methods runtime:", format(as.numeric(stop_time_resolution - start_time_resolution, units = "mins")), "minutes")
```


```{r}
sessionInfo()
```

